Hybrid nonlinear system identification using neural networks

### 2021-02-23 systemID_NNs.tex
working on the tensorflow script.
ok, working.
try to do with noise now.
ok, working on the empc script.
need to change the NEMPC classes to take funcs in discrete time.
add scripts.

A few alternative modeling choices are:
\begin{itemize}
	\item Black-box: $y^+ = f_N(y, z, u)$, (purely black-box)
  \item Hybrid 1: $x_g^+ = f_N(x_g, z, u), \ z^+ = f_z(z, x_g, u), \ y =
                   h_g(x_g)$, (states have meaning)
  \item Hybrid 2: $x_g^+ = f_g(x_g, u), \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g) +
                   h_N(z)$, (model residual)
  \item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)), \ z^+ = f_z(z, x_g,
  u), \ y = h_g(x_g)$, (use NNs to learn some parameters)
\end{itemize}

@@

The Figure \ref{fig:validation_pred} shows predictions of the trained black-box
and hybrid models after training with 8 hours (480 samples) of data. We observe
that both the models can match the plant data accurately. The black-box model
shows slightly worse predictions in a few places. However, both the trained
models are adequate for real-time optimization purpose in a feedback control system.

\begin{figure}[!h]
  \centering
  \includegraphics[page=2, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Model validation for the nonlinear reactor model. Both the black-box
           and hybrid model can predict the plant responses accurately. The
           model predictions shown in this plot correspond to training with 8
           hours of data.}
  \label{fig:validation_pred}
\end{figure}

We examine the prediction accuracy of the models as a function of the number of
training samples using the mean squared error (MSE) metric computed as follows
on the validation data
\begin{align*}
  \textnormal{MSE} =  \sum_{k=0}^{N_s-1} 
  \dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
\end{align*}
in which $\hat{y}$ are the model predictions and $y$ denotes the plant
measurements. The Figure \ref{fig:validation_mse} shows the variation of this
MSE metric as a function of the number of training data obtained with both the
black-box and hybrid models. We observe that the hybrid model provides better
performance than the black-box model in terms of overall prediction accuracy and
data requirements.

\begin{figure}[!h]
  \centering
  \includegraphics[page=9, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Mean squared error (MSE) performance metric obtained with the
           black-box and hybrid model with increase in the number of training
           samples. We notice that the hybrid model slightly outperforms the
           black-box model in data requirements and 
           overall prediction accuracy.}
  \label{fig:validation_mse}
\end{figure}

@@

\begin{align*}
  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
  x_s &= f(x_s, u_s)\\ 
  y_s &= h(x_s) \\
  \underline{u} \leq &u_s \leq \overline{u}
\end{align*}
in which $x_s$, $y_s$, $u_s$ are the state, output, and control respectively at
steady-states. For the nonlinear reactor example, the cost is chosen as
$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which measures the raw material cost and
product profit. The Figure \ref{fig:cost_curves} shows the cost curves of the
plant, grey-box, black-box, and hybrid model obtained with 8 hours of training
data. We first notice from the cost curve of the grey-box model that an
optimization with the grey-box model results in the optimizer recommending a
full supply of the raw material $A$. Since some of the product $B$ reacts to
form an undesired side product $C$ in the plant, this recommendation incurs a
cost as observed from the plant cost curve. An intuitive approach to correct
this type of plant-model mismatch is to augment the grey-box model with
integrating disturbance models and estimate the plant model mismatch from data
at one steady state. This strategy however, cannot recover the plant cost curve
in the entire actuator constraint set as the integrating disturbance models are
often linear. We notice that the cost curve of the hybrid model is almost the
same as the cost curve of the plant model in the entire actuator constraint set.
The cost curve of the black-box model is also shown and we notice that it is not
as closely aligned with the plant as compared to the hybrid model.

\begin{figure}[!h]
  \centering
  \includegraphics[page=6, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
  and hybrid model after training with 8 hours of data.}
  \label{fig:cost_curves}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[page=7, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Variation of the cost error metric ($e(C_{Af})$) in the actuator
           constraint set with increase in the training data for the 
           black-box neural network. The metric improves in the 
           constraint set with an increase in training data.}
  \label{fig:cost_error_bb}
\end{figure}

Finally, we analyze the variation in the cost curves of the black-box and hybrid
models with increasing number of training samples. We analyze the following cost
error metric in the entire actuator constraint set
\begin{align*}
  e(u_s) = \vert \ell(y_p(u_s), u_s) - \ell(y_m(u_s), u_s) \vert
\end{align*}
in which $y_p(\cdot)$ and $y_m(\cdot)$ denote the plant and model steady state
output respectively. The Figures \ref{fig:cost_error_bb} and
\ref{fig:cost_error_hybrid} show the variation in this cost error metric in the
actuator constraint set with increasing number of training samples for the
black-box and hybrid model respectively. We observe that with an increase in the
number of training samples, the error metric for the hybrid model approaches
near zero values for a wide range of actuator values in the constraint set. The
error metric of the black-box model remains relatively larger than that of the
hybrid model in the constraint set with increase in training data. The
implication of this error metric going to zero for all actuator values in the
constraint set is that the optimum computed for the steady-state economic
optimization problem with the trained model will be the same as the true plant
optimum.

\begin{figure}[!h]
  \centering
  \includegraphics[page=8, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Variation of the cost error metric ($e(C_{Af})$) in the actuator
  constraint set with increase in the training data for the 
  hybrid grey-box and neural network model. The metric improves in the 
  constraint set with an increase in training data.}
  \label{fig:cost_error_hybrid}
\end{figure}

\section{Large, CSTR and flash in series example}
\begin{figure}[!h]
  \centering
  \includegraphics[width=\textwidth, height=0.4\textheight]{cstr_flash.pdf}
  \caption{Schematic of the CSTR and flash plant. The CSTR facilitates two 
  reactions, the first reaction ($A \rightarrow B$) forms the primary product $B$, and the second reaction ($3B \rightarrow C$) 
  consumes the desired product $B$ and forms the undesired product $C$.
  The non-adiabatic flash separates the reactant $A$ and 
  products $B$ and $C$. The major component in the vapor phase 
  is $A$, and it is recycled back after cooling to the CSTR.}
  \label{fig:cstr_flash}
\end{figure}
\newpage
\textbf{Plant -- }
\begin{align*}
  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) + 
                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1 -3r_2\\
  \dfrac{dC_{Cr}}{dt} &= \dfrac{-FC_{Cr} + 
  D(C_{Cd} -C_{Cr})}{A_rH_r} + r_2\\
  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
                      \dfrac{r_1\Delta H_1 + r_2\Delta H_2}{\rho C_p} + 
                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
  \dfrac{dC_{Cb}}{dt} &= \dfrac{F_r(C_{Cr} -C_{Cb}) + 
                          D(C_{Cb} -C_{Cd})}{A_bH_b} \\
  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
\end{align*}

@@

The economic objective is -- 
\begin{align*}
  \underset{u_s}{\textnormal{min}} \quad 
  \ell(u_s, x_s) \quad := & \quad c_1F + c_2Q_r + 
                                      c_3Q_b - c_4C_{Bb} \\
  x_s &= f(x_s, u_s)\\ 
  y_s &= h(x_s) \\
  \underline{u} \leq &u_s \leq \overline{u}
\end{align*}

\end{document}

>>>>

A few alternative modeling choices are:
\begin{itemize}
  \item Black-box: $y^+ = f_N(y, z, u), \ z^+ = f_z(z, y, u)$
        (purely black-box).
  \item Hybrid 1: $x_g^+ = f_N(x_g, z, u), \ z^+ = f_z(z, x_g, u), \ y =
                   h_g(x_g)$, (states have meaning)
  \item Hybrid 2: $x_g^+ = f_g(x_g, u), \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g) +
                   h_N(z)$, (model residual)
  \item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)), \ z^+ = f_z(z, x_g,
  u), \ y = h_g(x_g)$, (use NNs to learn some parameters)
\end{itemize}

@@

The Figure shows predictions of the trained black-box
and hybrid models after training with 8 hours (480 samples) of data. We observe
that both the models can match the plant data accurately. The black-box model
shows slightly worse predictions in a few places. However, both the trained
models are adequate for real-time optimization purpose in a feedback control system.

We examine the prediction accuracy of the models as a function of the number of
training samples using the mean squared error (MSE) metric computed as follows
on the validation data
\begin{align*}
  \textnormal{MSE} =  \sum_{k=0}^{N_s-1} 
  \dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
\end{align*}
in which $\hat{y}$ are the model predictions and $y$ denotes the plant
measurements. The Figure  shows the variation of this
MSE metric as a function of the number of training data obtained with both the
black-box and hybrid models. We observe that the hybrid model provides better
performance than the black-box model in terms of overall prediction accuracy and
data requirements.

@@

\begin{align*}
  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
  x_s &= f(x_s, u_s)\\ 
  y_s &= h(x_s) \\
  \underline{u} \leq &u_s \leq \overline{u}
\end{align*}
in which $x_s$, $y_s$, $u_s$ are the state, output, and control respectively at
steady-states. For the nonlinear reactor example, the cost is chosen as
$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which measures the raw material cost and
product profit. The Figure shows the cost curves of the
plant, grey-box, black-box, and hybrid model obtained with 8 hours of training
data. We first notice from the cost curve of the grey-box model that an
optimization with the grey-box model results in the optimizer recommending a
full supply of the raw material $A$. Since some of the product $B$ reacts to
form an undesired side product $C$ in the plant, this recommendation incurs a
cost as observed from the plant cost curve. An intuitive approach to correct
this type of plant-model mismatch is to augment the grey-box model with
integrating disturbance models and estimate the plant model mismatch from data
at one steady state. This strategy however, cannot recover the plant cost curve
in the entire actuator constraint set as the integrating disturbance models are
often linear. We notice that the cost curve of the hybrid model is almost the
same as the cost curve of the plant model in the entire actuator constraint set.
The cost curve of the black-box model is also shown and we notice that it is not
as closely aligned with the plant as compared to the hybrid model.

Finally, we analyze the variation in the cost curves of the black-box and hybrid
models with increasing number of training samples. We analyze the following cost
error metric in the entire actuator constraint set
\begin{align*}
  e(u_s) = \vert \ell(y_p(u_s), u_s) - \ell(y_m(u_s), u_s) \vert
\end{align*}
in which $y_p(\cdot)$ and $y_m(\cdot)$ denote the plant and model steady state
output respectively. The Figures and show the variation in this cost error metric in the
actuator constraint set with increasing number of training samples for the
black-box and hybrid model respectively. We observe that with an increase in the
number of training samples, the error metric for the hybrid model approaches
near zero values for a wide range of actuator values in the constraint set. The
error metric of the black-box model remains relatively larger than that of the
hybrid model in the constraint set with increase in training data. The
implication of this error metric going to zero for all actuator values in the
constraint set is that the optimum computed for the steady-state economic
optimization problem with the trained model will be the same as the true plant
optimum.

\section{Large, CSTR and flash in series example}
\begin{figure}[!h]
  \centering
  \includegraphics[width=\textwidth, height=0.4\textheight]{cstr_flash.pdf}
  \caption{Schematic of the CSTR and flash plant. The CSTR facilitates two 
  reactions, the first reaction ($A \rightarrow B$) forms the primary product $B$, and the second reaction ($3B \rightarrow C$) 
  consumes the desired product $B$ and forms the undesired product $C$.
  The non-adiabatic flash separates the reactant $A$ and 
  products $B$ and $C$. The major component in the vapor phase 
  is $A$, and it is recycled back after cooling to the CSTR.}
  \label{fig:cstr_flash}
\end{figure}
%\newpage
\textbf{Plant -- }
\begin{align*}
  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) +
                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1 -3r_2\\
  \dfrac{dC_{Cr}}{dt} &= \dfrac{-FC_{Cr} + 
  D(C_{Cd} -C_{Cr})}{A_rH_r} + r_2\\
  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
                      \dfrac{r_1\Delta H_1 + r_2\Delta H_2}{\rho C_p} + 
                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
\end{align*}

\begin{align*}
  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
  \dfrac{dC_{Cb}}{dt} &= \dfrac{F_r(C_{Cr} -C_{Cb}) + 
                          D(C_{Cb} -C_{Cd})}{A_bH_b} \\
  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
\end{align*}

@@

The economic objective is -- 
\begin{align*}
  \underset{u_s}{\textnormal{min}} \quad 
  \ell(u_s, x_s) \quad := & \quad c_1FC_{Af} + c_2Q_r + 
                                      c_2Q_b + c_2D\rho C_p(T_d-T_b) - 
                                      c_3F_bC_{Bb} \\
  x_s &= f(x_s, u_s)\\ 
  y_s &= h(x_s) \\
  \underline{u} \leq &u_s \leq \overline{u}
\end{align*}

in which $c_1$ is the raw material cost of unit $\$$/mol-A,
$c_2$ is the energy price of unit $\$$/kW, 
and $c_3$ is the product selling price of unit $\$$/mol-B.
\end{document}
